---
title: "MA 615 Strawberry EDA"
author: "Chenxuan Xiong"
format: html
editor: visual
---

## Data acquisition and assessment

[Data Source](https://quickstats.nass.usda.gov/) [The data have been stored on NASS here:](https://quickstats.nass.usda.gov/results/45FBC825-B104-38E2-9802-839F5F3C7036)

Original Conditions:

Organization: The dataset is consists of basic information of strawberry growing information include year, period, price, state and organic conditions etc.

Problems: There are lots of NAs and single value columns. Also there exists missing values in Value and CV filled with (D), (H) or just blank cells. Records in Data.item column is messy. - Data sources\
- Assumptions and motivations

## Data cleaning and organization

```{r}
#| label: Load libraries
#| warning: false
#| message: false
#| echo: false

library(knitr)  
library(kableExtra)
library(tidyverse)
library(stringr)
```

Step 1: Remove columns with only NAs.

```{r}
#| label: drop one-item columns
#| echo: false
#read data
strawberry <- read.csv("strawberry.csv")
## define function
#drop NA
strawberry <- strawberry %>%
  select_if(~ !all(is.na(.)))
```

Step 2: Remove columns with only single value.

```{r}
#drop single value columns
single_value <- function(vec){
  tab <- table(vec)
  if (length(tab) == 1) {
    return(TRUE)
  }
  else{
    return(FALSE)
  }
}


drop_single_value_col <- function(df) {
  cols_to_drop <- c()
  for (i in 1:(ncol(df) - 1)) {
    if (single_value(df[, i])) {
      cols_to_drop <- c(cols_to_drop, i)
    }
  }
  df <- df[, -cols_to_drop]
  return(df)
}

strawberry1 <- drop_single_value_col(strawberry)


strawberry1$Value <- as.numeric(strawberry1$Value)
strawberry_census <- subset(strawberry1, Program == "CENSUS")
strawberry_survey <- subset(strawberry1, Program == "SURVEY")
View(strawberry)
```

Step 3: Deal with the (D) value in the Value column. We replace (D) values by the mean of "Value" of each data item group.

```{r}
##First we drop all (D)s ,(Z)s and (NA)s in the dataset.
strawberry2 <- strawberry1[complete.cases(strawberry1), ]
strawberry1$Value <- as.numeric(strawberry1$Value)
strawberry2 <- strawberry1[complete.cases(strawberry1), ]
uniq_data_item <- unique(strawberry2$Data.Item)
#mean_value by data item
mean_value <- strawberry2 |>
  group_by(Data.Item) |>
  summarize(MeanValue = mean(Value))
mean_value <- as.data.frame(mean_value)

#replace value by mean
strawberry_census_1 <- strawberry_census %>%
  group_by(Data.Item) %>%
  mutate(Value = ifelse(is.na(Value), mean(Value, na.rm = TRUE), Value))


strawberry_survey_1 <- strawberry_survey %>%
  group_by(Data.Item) %>%
  mutate(Value = ifelse(is.na(Value), mean(Value, na.rm = TRUE), Value))


strawberry_survey_1 <- strawberry_survey_1 %>%
  mutate(
    Chemical = if_else(str_detect(Domain.Category, "\\(.*=.*\\)"),
                       str_extract(Domain.Category, "(?<=\\().*?(?=\\=)"),
                       NA_character_),
    Chemical_Code = if_else(str_detect(Domain.Category, "\\(.*=.*\\)"),
                            str_extract(Domain.Category, "(?<=\\=).*?(?=\\))"),
                            NA_character_)
  )



# strawberry_survey_1$Chemical_Code_num <- as.numeric(strawberry_survey_1$Chemical_Code)
# strawberry_survey_1$Chemical_Code_str <- ifelse(is.na(strawberry_survey_1$Chemical_Code_num),
#                                         NA,
#                                         sprintf("%06d", strawberry_survey_1$Chemical_Code_num))
# library(httr)
# library(jsonlite)
# get_cas <- function(PC){
#     path <- paste0("https://ordspub.epa.gov/ords/pesticides/apprilapi/?q=%7b%22ais%22:%7b%22$instr%22:%22", PC,"%22%7d%7d")
#     r <- GET(url = path)
#     r_text <- content(r, as = "text", encoding = "UTF-8")
#     df <- fromJSON(r_text, flatten = TRUE)
#     df_strwb <- df$items[grepl("Strawberries", df$items$sites, fixed=T),]
#     ais <- df_strwb$ais[1]
#     pattern <- "\\(([^A-Za-z]+)\\/([0-9-]+)\\)"
#     text <- ais
#     matches <- regmatches(text, gregexpr(pattern, text))
#     cas <- sapply(matches, function(x) gsub(".*\\/([0-9-]+)\\)", "\\1", x)) 
#     if (is.character(cas)) {
#         return(cas[1])
# } 
#     else {
#         return("can't find")
# }
# }
# unique_stb=unique(strawberry_survey_1$Chemical_Code_str)
# result=numeric()
# k=numeric()
# for(i in 1:length(unique_stb)){
#   result[i]=get_cas(unique_stb[i])
#   k[i]=unique_stb[i]
#   print(result[i])
# }
# 
# code_dict <- data.frame("pc_code" = k,
#                           "cas_code" = result)
write_csv(code_dict,"code_dict.csv")
code_dict <- read.csv("code_dict.csv")


WHO_CAS_to_SeverityTable <- read_excel("WHO - CAS-to-SeverityTable.xlsx")

code_dict$Chemical_Code_num <- as.numeric(code_dict$k)
code_dict$Chemical_Code_str <- ifelse(is.na(code_dict$Chemical_Code_num),
                                         NA,
                                         sprintf("%06d", code_dict$Chemical_Code_num))
cas <- read.csv("CAS.csv",header=T)
code_dict$pc_code <- as.numeric(code_dict$pc_code)
code_dict$pc_code <- ifelse(is.na(code_dict$pc_code),NA,sprintf("%06d", code_dict$pc_code))
merged_df_all <- read.csv("merage(1).csv")

merged_df_all$Chemical_Code_num <- as.numeric(merged_df_all$k)
merged_df_all$Chemical_Code_str <- ifelse(is.na(merged_df_all$Chemical_Code_num),
                                         NA,
                                         sprintf("%06d", merged_df_all$Chemical_Code_num))

merged_df_strw <- merge(merged_df_all, strawberry_survey_1,by = "Chemical_Code_str")
write_csv(merged_df_strw,"merged_df_strw.csv")
```

Step 4: We find the severity of chemical used by converting PC code in the data set to CAS code. And then find the corresponding pesticide and its toxicity.

```{r}

strawberry_survey_1 <- strawberry_survey_1 %>%
  mutate(
    Chemical = if_else(str_detect(Domain.Category, "\\(.*=.*\\)"),
                       str_extract(Domain.Category, "(?<=\\().*?(?=\\=)"),
                       NA_character_),
    Chemical_Code = if_else(str_detect(Domain.Category, "\\(.*=.*\\)"),
                            str_extract(Domain.Category, "(?<=\\=).*?(?=\\))"),
                            NA_character_)
  )



# strawberry_survey_1$Chemical_Code_num <- as.numeric(strawberry_survey_1$Chemical_Code)
# strawberry_survey_1$Chemical_Code_str <- ifelse(is.na(strawberry_survey_1$Chemical_Code_num),
#                                         NA,
#                                         sprintf("%06d", strawberry_survey_1$Chemical_Code_num))
# library(httr)
# library(jsonlite)
# get_cas <- function(PC){
#     path <- paste0("https://ordspub.epa.gov/ords/pesticides/apprilapi/?q=%7b%22ais%22:%7b%22$instr%22:%22", PC,"%22%7d%7d")
#     r <- GET(url = path)
#     r_text <- content(r, as = "text", encoding = "UTF-8")
#     df <- fromJSON(r_text, flatten = TRUE)
#     df_strwb <- df$items[grepl("Strawberries", df$items$sites, fixed=T),]
#     ais <- df_strwb$ais[1]
#     pattern <- "\\(([^A-Za-z]+)\\/([0-9-]+)\\)"
#     text <- ais
#     matches <- regmatches(text, gregexpr(pattern, text))
#     cas <- sapply(matches, function(x) gsub(".*\\/([0-9-]+)\\)", "\\1", x)) 
#     if (is.character(cas)) {
#         return(cas[1])
# } 
#     else {
#         return("can't find")
# }
# }
# unique_stb=unique(strawberry_survey_1$Chemical_Code_str)
# result=numeric()
# k=numeric()
# for(i in 1:length(unique_stb)){
#   result[i]=get_cas(unique_stb[i])
#   k[i]=unique_stb[i]
#   print(result[i])
# }
# 
# code_dict <- data.frame("pc_code" = k,
#                           "cas_code" = result)
write_csv(code_dict,"code_dict.csv")
code_dict <- read.csv("code_dict.csv")


WHO_CAS_to_SeverityTable <- read_excel("WHO - CAS-to-SeverityTable.xlsx")

code_dict$Chemical_Code_num <- as.numeric(code_dict$k)
code_dict$Chemical_Code_str <- ifelse(is.na(code_dict$Chemical_Code_num),
                                         NA,
                                         sprintf("%06d", code_dict$Chemical_Code_num))
cas <- read.csv("CAS.csv",header=T)
code_dict$pc_code <- as.numeric(code_dict$pc_code)
code_dict$pc_code <- ifelse(is.na(code_dict$pc_code),NA,sprintf("%06d", code_dict$pc_code))
merged_df_all <- read.csv("merage(1).csv")

merged_df_all$Chemical_Code_num <- as.numeric(merged_df_all$k)
merged_df_all$Chemical_Code_str <- ifelse(is.na(merged_df_all$Chemical_Code_num),
                                         NA,
                                         sprintf("%06d", merged_df_all$Chemical_Code_num))

merged_df_strw <- merge(merged_df_all, strawberry_survey_1,by = "Chemical_Code_str")
write_csv(merged_df_strw,"merged_df_strw.csv")

```

## EDA

### Severity level distribution of pesticides in each year

```{r}

```

### Severity level distribution of pesticides used in each state

```{r}

```

### The most popular kind of chemical by year

```{r}

```

### The most popular kind of chemical by state

```{r}

```

## References

### Material about strawberries

[WHO says strawberries may not be so safe for you--2017March16](https://med.news.am/eng/news/13621/who-says-strawberries-may-not-be-so-safe-for-you.html)

[Pesticides + poison gases = cheap, year-round strawberries 2019March20](https://www.ewg.org/foodnews/strawberries.php)

[Multistate Outbreak of Hepatitis A Virus Infections Linked to Fresh Organic Strawberries-2022March5](https://www.cdc.gov/hepatitis/outbreaks/2022/hav-contaminated-food/index.htm)

[Strawberry makes list of cancer-fighting foods-2023May31](https://issuu.com/mechlocal/docs/053123_mech_asf/s/25386339)

### Technical references

In their handbook ["An introduction to data cleaning with R" by Edwin de Jonge and Mark van der Loo](https://cran.r-project.org/doc/contrib/de_Jonge+van_der_Loo-Introduction_to_data_cleaning_with_R.pdf), de Jonge and van der Loo go into detail about specific data cleaning isssues and how to handle them in R.

["Problems, Methods, and Challenges in Comprehensive Data Cleansing" by Heiko Müller and Johann-Christoph Freytag](https://www.researchgate.net/profile/Heiko-Mueller/publication/228929938_Problems_methods_and_challenges_in_comprehensive_data_cleansing/links/09e415101b58541e2c000000/Problems-methods-and-challenges-in-comprehensive-data-cleansing.pdf) is a good companion to the de Jonge and van der Loo handbook, offering additional insights.

## Initial questions

## The data

[Data Source](https://quickstats.nass.usda.gov/) [The data have been stored on NASS here:](https://quickstats.nass.usda.gov/results/45FBC825-B104-38E2-9802-839F5F3C7036)

Original Conditions:

Organization:

Problems:There are lots of NAs and single value columns. Also there exists missing values in Value and CV filled with (D), (H) or just blank cells. Records in Data.item column is messy.
